---
title: "EDA"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(haven)
```

Import the weather data.

```{r}
weather_df = read_csv("./data/weather_df.csv")
weather_df = weather_df |>
  mutate(
    month = floor_date(date, unit = "month")) |>
  select(name, id, everything())
```

## group_by()

```{r}
weather_df |> 
  group_by(name, month)
```
可以用ungroup()取消操作。


counting stuff：
n_distinct() only count uniqe ():
```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(n_obs = n())

weather_df |> 
  group_by(month) |> 
  summarize(n_obs = n(), n_days = n_distinct(date))
```
不用group by，和上面一样的结果：
```{r}
weather_df |> 
  count(name, month, name = "n_obs")
```



## 2x2 
NEVER ues table function.
cold = case_when(
      tmax < 5 ~ "cold",
      tmax >=5 ~ "not_cold"
    )当tmax < 5时，cold=“cold”，用同一列表示：
```{r}
weather_df |> 
  drop_na(tmax) |> 
  filter(name != "Molokai_HI") |> 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >=5 ~ "not_cold"
    )
  ) |> 
  group_by(name, cold) |> 
  summarize(count = n())
```

2x2 table：
```{r}
weather_df |> 
  drop_na(tmax) |> 
  filter(name != "Molokai_HI") |> 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >=5 ~ "not_cold"
    )
  ) |> 
  janitor::tabyl(name, cold)
```





## general numeric summaries.

let's try some other useful summaries.
na.rm = TRUE去掉NA：

```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE),
    sd_prcp = sd(prcp, na.rm = TRUE)
  )
```
简化上面的mean步骤,同时mean prcp，tmax，tmin：
```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(across(prcp:tmin, mean)
  )
```

summarize and then plot ...
```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE),
    sd_prcp = sd(prcp, na.rm = TRUE) #same as above#
  ) |> 
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
  geom_point() + 
  geom_line()
```


easier format for readers to read:
knitr::kable(
    digits = 3, 保留3位小数
    col.names = c("Month", "Central Park", "Molokai", "Waterhole")改名字：
```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) |> 
  pivot_wider(
    names_from = name, 
    values_from = mean_tmax
  ) |> 
  knitr::kable(
    digits = 3, 
    col.names = c("Month", "Central Park", "Molokai", "Waterhole"))
```


## grouped mutates

```{r}
weather_df |> 
  group_by(name) |> 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax) |> 
  ggplot(aes(x = date, y = centered_tmax, color = name)) +
  geom_point()
```

Find hottest / coldest days.
min_rank（）从小到大排列；
 min_rank(desc(））从大到小排列：
```{r}
weather_df |> 
  group_by(name) |> 
  mutate(
    temp_rank = min_rank(desc(tmax))
  ) |> 
  filter(temp_rank < 4)


weather_df |> 
  group_by(name) |> 
  filter(min_rank(tmax) < 4) |> 
  arrange(tmax)
```

lag(tmax, 5)把所有tmax往下移5行：
```{r}
weather_df |> 
  group_by(name) |> 
  mutate(
    lagged_tmax = lag(tmax),
    temp_change = tmax - lagged_tmax
  ) |> 
  filter(min_rank(temp_change) < 3)
```



```{r}
weather_df |> 
  group_by(name) |> 
  mutate(
    lagged_tmax = lag(tmax),
    temp_change = tmax - lagged_tmax
  ) |> 
  summarize(
    sd_tmax_change = sd(temp_change, na.rm = TRUE)
  )
```





