---
title: "Visualization"
subtitle: "part 2"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
library(haven)
```

Import the weather data

```{r}
weather_df = read_csv("./data/weather_df.csv")
weather_df
```


Make a scatterplot but fancy this time.
labs()加label
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs(
    title = "Temperature scatterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location", 
    caption = "Weather data taken from rnoaa package for three stations."
  )
```

Scales -- start with `x` and `y` and then do `color`.
scale_x_continuous(), scale_y_continuous()控制横竖坐标的显示；
transform = "sqrt"坐标是按照square root表示的；同理transform = "log"；
position = "right"坐标在图的右边：
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs(
    title = "Temperature scatterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location", 
    caption = "Weather data taken from rnoaa package for three stations."
  ) + 
  scale_x_continuous(
    breaks = c(-15, 0, 20),
    labels = c("-15C", "0", "20")
  ) + 
  scale_y_continuous(
    limits = c(0, 30),
    transform = "sqrt",
    position = "right"
  )
```


Look at color:
scale_color_viridi()默认是continuous variables
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs(
    title = "Temperature scatterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location", 
    caption = "Weather data taken from rnoaa package for three stations."
  ) + 
  viridis::scale_color_viridis(discrete = TRUE)
```



## Themes

讲上面的图保存为ggp_scatterplot：
```{r}
ggp_scatterplot = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs(
    title = "Temperature scatterplot",
    x = "Minimum Temp (C)",
    y = "Maximum Temp (C)",
    color = "Location", 
    caption = "Weather data taken from rnoaa package for three stations."
  ) + 
  viridis::scale_color_viridis(discrete = TRUE)

```
将图例放在最下面： 
```{r}
ggp_scatterplot + 
  theme(legend.position = "bottom")
```
theme_bw()将底色变白：
```{r}
ggp_scatterplot + 
  theme_bw() + 
  theme(legend.position = "bottom")
```

Order matters ...

```{r}
ggp_scatterplot + 
  theme(legend.position = "bottom") + 
  theme_bw() 
```
theme_minimal()将最外面的黑框去掉：
```{r}
ggp_scatterplot + 
  theme(legend.position = "bottom") + 
  theme_minimal() 
```

theme_classic()将网格去掉：
```{r}
ggp_scatterplot + 
  theme(legend.position = "bottom") + 
  theme_classic() 
```

ggthemes::theme_excel() looks like excel2003:
```{r}
ggp_scatterplot + 
  theme(legend.position = "bottom") + 
  ggthemes::theme_excel() 
```
### set options
```{r}
library(tidyverse)

knitr::opts_chunk$set(fig.width = 6, fig.asp = .6, output_width = "90%")

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(ggplot2.continuous.colour = "viridis",
        ggplot2.continuous.fill = "viridis")

scale_color_discrete = scale_color_viridis_d()
scale_fill_discrete = scale_fill_viridis_d()

central_park_df = 
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai_df = 
  weather_df |> 
  filter(name == "Molokai_HI")
```

一个图中用两组dataset：
```{r}
ggplot(data = molokai_df, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park_df) 
```


### patchwork
(tmax_tmin_p + prcp_dens_p) / tmax_date_p将三个图放在一起，前两个在第一行，/后面的在第二行
```{r}
tmax_tmin_p = 
  weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```


### Data Manipulation
为了更改x轴分类的顺序，使得按照特别的名字顺序排列，forcats::fct_relevel（），所以按照"Molokai_HI", "CentralPark_NY", "Waterhole_WA"的顺序排列：
```{r}
weather_df |>
  mutate(name = forcats::fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |> 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```
同时表现tmin和tmax：
```{r}
weather_df |>
  select(name, tmax, tmin) |> 
  pivot_longer(
    tmax:tmin,
    names_to = "observation", 
    values_to = "temp") |> 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) + 
  facet_grid(~name) + 
  viridis::scale_fill_viridis(discrete = TRUE)
```




