---
title: "Viz I"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)

```

Import the weather data

```{r}
weather_df = read_csv("./data/weather_df.csv")
weather_df
```

Making our first plot : geom_point()用来把点给加上


```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

和上面一样的图：
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

save a plot object:
```{r}
ggp_weather_scatterplot = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

ggp_weather_scatterplot
```

Check why some rows are missing.
把所有na筛选出来：
```{r}
weather_df |> 
  filter(is.na(tmax))
```


## Fancier scatterplots!
geom_point加点，geom_smooth划线：
color = name also aply to geom_smooth(),则分组划线；
alpha = .3是每个点有30%的透明度；
size = .8是每个点的大小：
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3, size = .8) + 
  geom_smooth(se = FALSE)
```

Where you define aesthetics can matter

color only apply to geom_point()，则不分组划线
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .3, size = .8) + 
  geom_smooth(se = FALSE)
```

use faceting real quick

facet_grid(. ~ name)不同name作为一个column各画一个图，3个竖着的图；
facet_grid(name ~ .)不同name作为一个row各画一个图，3个横着的图;
alpha = tmin不同的tmin有不同的透明度：
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, alpha = tmin, color = name)) + 
  geom_point() + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

Let's make a somewhat more interesting scatterplot

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) + 
  geom_point(alpha = .3) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```


LA plot:

how many geoms have to exist?
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_smooth(se = FALSE)

```


## Small things

geom_hex()蜂窝状，geom_density2d()等高线：
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex()

weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_density2d()
```


```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(color = "blue")
```

## Univariate plots
histogram：

```{r}
weather_df |> 
  ggplot(aes(x = tmin, color = name)) + 
  geom_histogram()
```
position = "dodge" 不同分组并排；
fill = name没有框：
```{r}
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_histogram(position = "dodge")
```

how would i fix this? maybe facet?
fill = name加上颜色：
```{r}
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_histogram() + 
  facet_grid(. ~ name)
```

maybe a density plot?
geom_density()平滑的histogram：
```{r}
weather_df |> 
  ggplot(aes(x = tmin, fill = name)) + 
  geom_density(alpha = .3)
```
boxplot：
```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmin, fill = name)) + 
  geom_boxplot()
```

violin plots平滑的box plot;
stat_summary(fun = "median") show median point:
```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmin, fill = name)) + 
  geom_violin() +
  stat_summary(fun = "median")
```

ridge plot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges(alpha = .5) 
```




## Saving and embedding plots

saving plots

```{r}
ggp_weather = 
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()


ggsave("plots/ggp_weather.pdf", ggp_weather, width = 8, height = 6)
```


embedding plots

```{r, fig.width = 6, fig.hight = .6}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()
```
```{r, fig.width = 12, fig.hight = 6}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()
```





